---

title: "Introduction"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Introduction}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
library(MuniCandS)
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```


# Introduction
The MuniCandS package provides a powerful framework for testing goodness-of-fit for multivariate distributions. Unlike traditional methods that rely on a single global statistic, MuniCandS uses a decomposition of the empirical process into asymptotically independent Gaussian components.

This approach, based on the work of Cabaña & Cabaña (2025), allows for more nuanced testing (the m-test and s-test) across various distributional families.

# Methodology
The methodology for testing uniformity on the hypercube is based on a decomposition of finite signed measures into zero-marginal components, leading to a decomposition of the empirical process into asymptotically independent terms.

The resulting procedures are of Cramér–von Mises type, but replace the quadratic norm of the empirical process by the collection of squared norms of its zero-marginal components.

The fact that whenever a null hypothesis implies a joint distribution that factorizes into independent continuous components after a suitable mapping, allows to test such hypotheses by reducing the problem to a uniformity test on the hypercube via componentwise probability integral transforms.

Specifically, the package introduces new procedures derived from these principles to implement multivariate tests of:

# Supported hypotheses

| Code   | Hypothesis                    |
| ------ | ----------------------------- |
| `"UC"` | Uniformity on the hypercube   |
| `"US"` | Uniformity on the hypersphere |
| `"N"`  | Multivariate normality        |
| `"I"`  | Spherical symmetry            |
| `"E"`  | Elliptical symmetry           |
| `"IN"` | Independence                  |


# Quick Start: Testing Normality

Let's test if a 3-dimensional sample comes from a standard multivariate normal distribution.

```{r}
set.seed(42)
# Generate a sample under H0
n <- 100
p <- 3
X <- matrix(rnorm(n * p), nrow = n, ncol = p)

# Run the test
results <- MuniCandS(X, type = "N", n_sim = 100, n_mc = 100)

# Display p-values
print(results$result)
```

# Efficient Workflow: Using the Cache

 The most computationally intensive part of MuniCandS is the Monte Carlo simulation of the null distributions. These depend on the dimensions (n, p) and parameters, but not on the data itself.

### Step 1: Generate the Cache

 ```{r cache_gen}
 res_first <- MuniCandS(
   X, 
   type = "N", 
   n_sim = 100, 
   n_mc = 100, 
   return_cache = TRUE
 )
 ```

### Step 2: Reuse for New Data

 ```{r cache_reuse}
 X_new <- matrix(rnorm(300), ncol = 3)

 res_second <- MuniCandS(
   X_new, 
   type = "N", 
   n_sim = 100,
   n_mc = 100,
   cache = res_first$cache
 )
 ```

# Interpreting Results

 The MuniCandS function returns two main p-values:

*   m-test: Based on the minimum of the components' p-values.
*   s-test: Based on a sum of functions of the p-values.

 A small p-value (typically < 0.05) in either test suggests evidence against the null hypothesis.

# Mathematical Background

 For more details, please refer to:

 Cabaña, A. and Cabaña, E. M. (2025). *Brownian sheet and uniformity tests on the hypercube*. Statistica (to appear). arXiv:2509.06134.


