---

title: "Introduction to MuniCandS"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Introduction to MuniCandS}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```


# Introduction

`MuniCandS` implements multivariate tests of:

* uniformity on the hypercube,
* uniformity on the hypersphere,
* normality,
* spherical symmetry,
* elliptical symmetry,
* independence.

The methodology for testing uniformity on the hypercube is based on a decomposition of finite signed measures into zero-marginal components, leading to a decomposition of the empirical process into asymptotically independent terms.

The resulting procedures are of Cramér–von Mises type, but replace the quadratic norm of the empirical process by the collection of norms of its zero-marginal components.

Further theoretical details can be found in:

A. Cabaña and E. M. Cabaña (2025),
*Brownian sheet and uniformity tests on the hypercube*,
Statistica, to appear,
arXiv:2509.06134.


The fact that whenever a null hypothesis implies a joint distribution that factorizes into independent continuous components after a suitable mapping, allows to test such hypotheses by reducing the problem to a uniformity test on the hypercube via componentwise probability integral transforms.
Specifically, the package introduces new procedures derived from these principles for testing uniformity on the hypersphere $S^p$, as well as multivariate normality, spherical and elliptical symmetry, and independence in $R^p$.



# Basic example

```{r}
library(MuniCandS)

set.seed(1)

Z <- matrix(runif(200), ncol = 2)

MuniCandS(
  Z,
  type = "UC",
  n_sim = 100,
  n_mc = 100
)
```

# Supported hypotheses

| Code   | Hypothesis                    |
| ------ | ----------------------------- |
| `"UC"` | Uniformity on the hypercube   |
| `"US"` | Uniformity on the hypersphere |
| `"N"`  | Multivariate normality        |
| `"I"`  | Spherical symmetry            |
| `"E"`  | Elliptical symmetry           |
| `"IN"` | Independence                  |

# Monte Carlo reuse

The function supports optional caching of Monte Carlo simulations:

```{r}
res <- MuniCandS(
  Z,
  type = "UC",
  n_sim = 100,
  n_mc = 100,
  return_cache = TRUE
)

cache <- res$cache
```
The cached Monte Carlo simulations can then be reused for another sample of the same dimension and test configuration:

```{r}
Z2 <- matrix(runif(200), ncol = 2)

MuniCandS(
  Z2,
  type = "UC",
  n_sim = 100,
  n_mc = 100,
  cache = cache
)
```

